Baseline centering problem

### 2018-12-12 paper/tangotd.tex

Let us consider the norm
\begin{equation}
\norm{\Delta V-R}^2_{P,\tilde P}\deq \transp{(\tfrac12 (\Delta+\tilde
\Delta)
V-R)}\rho(\tfrac12 (\Delta+\tilde \Delta)V-R)
\end{equation}
which is the $L^2(\rho)$-norm of the Bellman gap of an environment
$\frac12 (P+\tilde P)$ in which half of the transitions are real or
simulated.

The gradient of this norm is
\begin{equation}
\tfrac12 \transp{\Phi}\transp{(\Delta+\tilde
\Delta)
}\rho((\Delta+\tilde \Delta)V-R)
\end{equation}

We can easily compute
\begin{equation}
\tfrac12 \transp{\Phi}\transp{\Delta}
\rho(\tilde \Delta V-R)
+
\tfrac12 \transp{\Phi}\transp{\tilde \Delta}
\rho(\Delta V-R)
\end{equation}

>>>>

Therefore, by various combinations of actual and simulated transitions,
we can easily compute samples of
\begin{equation}
\tfrac12 \transp{\Phi}(\transp{\Delta}\rho \tilde \Delta+\transp{\tilde
\Delta}\rho\Delta)V
-
\transp{\Phi}\transp{\tilde \Delta}
\rho R
\end{equation}
(variant 1) and
\begin{equation}
\tfrac12 \transp{\Phi}\transp{\Delta}
\rho(\tilde \Delta V-R)
+
\tfrac12 \transp{\Phi}\transp{\tilde \Delta}
\rho(\Delta V-R)
\end{equation}
(variant 2, under the conditions of the footnote).
Setting
\begin{equation}
C\deq \tfrac12 (\transp{\Delta}\rho \tilde \Delta+\transp{\tilde
\Delta}\rho\Delta)
\end{equation}
variant 1 is the gradient of
\begin{equation}
\norm{V-C^{-1}\transp{\tilde\Delta}\rho R}^2_C
\end{equation}
while variant 2 is the gradient of
\begin{equation}
\norm{V-\frac12 C^{-1}(\transp{\tilde\Delta}+\transp{\Delta})\rho R}^2_C
\end{equation}

Therefore, if $C$ is definite positive, then variant 1 and variant 2 will converge to
the best approximation (in the parametric family $V$) of
\begin{equation}
\hat V_1\deq C^{-1}\transp{\tilde\Delta}\rho R \qquad\text{and}\qquad \hat
V_2\deq \frac12
C^{-1}(\transp{\tilde\Delta}+\transp{\Delta})\rho R
\end{equation}
respectively.
If $\tilde \Delta=\Delta$, then both $\hat V_1$ and $\hat V_2$
are equal to the desired target
$\Delta^{-1}R$.

Let us study how these fixed points approximate the true target
$\Delta^{-1}R$ when $\tilde \Delta \approx \Delta$. First,
\begin{equation}
C=\frac14 \transp{(\tilde \Delta+\Delta)}\rho(\tilde \Delta+\Delta)-\frac14
\transp{(\tilde
\Delta-\Delta)}\rho(\tilde
\Delta-\Delta)
\end{equation}
Therefore, $C$ is definite positive provided
\begin{equation}
\transp{(\tilde
\Delta-\Delta)}\rho(\tilde
\Delta-\Delta)\prec \transp{(\tilde \Delta+\Delta)}\rho(\tilde \Delta+\Delta)
\end{equation}
which is the case when the model $\tilde \Delta$ is close enough
to the true environment $\Delta$ (moreover the left-hand-side is at
second order in the error). Therefore, convergence is guaranteed in a relatively large
zone around $\Delta$. But the method will diverge if $\tilde \Delta$ is
too far from $\Delta$—contrary to double sampling using the model twice,
which always converges to the value function of the model.

Next, since $C=\frac14 \transp{(\tilde \Delta+\Delta)}\rho(\tilde
\Delta+\Delta)+O(\tilde\Delta-\Delta)^2$, the fixed point $\hat V_2$ for variant 2
is
\begin{equation}
\hat V_2=\left(\frac{\tilde
\Delta+\Delta}{2}\right)^{-1}R+O(\tilde\Delta-\Delta)^2
\end{equation}
namely, at first order, the fixed point of variant 2 is the value
function of a Markov chain with transitions $\tfrac12(\tilde P+P)$,
namely, each transition is real or simulated with probability $1/2$.
Intuitively this should keep reasonable properties of the true value
function: even if the model $\tilde P$ diverges when iterated, the Markov chain $\tfrac12(\tilde
P+P)$ will have a dynamics closer to the true environment.
